### 2020-02-14 CalNotes.tex
Proofreading

\textbf{Abstract.} We study the problem of drift estimation for two-scale continuous time series. We set ourselves in the framework of overdamped Langevin equations, for which the existence of a single-scale surrogate homogenized equation exists. In this setting, estimating the drift coefficient of the homogenized equation requires pre-processing of the data, often in the form of subsampling. We avoid subsampling by filtering the data with an appropriate kernel function and computing maximum likelihood estimators based on the filtered process. We show that the estimators we propose are asymptotically unbiased and demonstrate numerically the advantages of our method with respect to subsampling. Finally, we show how our filtering methodology can be combined with Bayesian techniques and provide a full uncertainty quantification of the inference procedure.

@@

Data-driven approaches to the derivation of effective stochastic coarse-grained dynamics has become a standard tool in different areas such as molecular dynamics or atmosphere and ocean science \cite{KKP15,CoP09,YMV19}. When fitting data to a low dimensional coarse-grained model, one is faced with the problem of model misspecification; indeed, the data, coming from the full dynamics, is compatible with the coarse-grained model only at the time scales at which the effective dynamics is valid. This is related to the issue of high-frequency data that arises in econometrics \cite{AiJ14,AMZ05,ZMA05,OSP10}.

The issue of model misspecification in inverse problems with a multiscale structure has been treated in the context of partial differential equations (PDE). In particular, it has been shown that it is possible to infer a coarse-grained equation from data coming from the full model and to retrieve asymptotically the correct result \cite{NPS12}. A series of papers \cite{AbD18, AbD19, AGZ19} focuses on retrieving the full model when the multiscale coefficient is endowed with a specific parametrized structure. Being these problems ill-posed, the latter is achieved via Tikhonov regularization \cite{AbD19,NPS12}, adopting a Bayesian approach \cite{AbD18, NPS12} or exploiting techniques of Kalman filtering \cite{AGZ19}. In \cite{AbD18,AGZ19}, the authors highlight the need of accounting explicitly for the modelling error due to homogenisation and apply statistical techniques taken from \cite{CDS18,CES14}.

For simple models in molecular dynamics, the effect of model misspecification was studied in a series of papers \cite{PaS07,PPS09,PPS12,GaS17,GaS18} under the assumption of scale separation. In particular, for Brownian particles moving in two-scale potentials it was shown that, when fitting data from the full dynamics to the homogenised equation, the maximum likelihood estimator (MLE) is asymptotically biased \cite[Theorem 3.4]{PaS07}. To be more precise, in the large sample size limit the MLE converges to the coefficients of the unhomogenised equation, rather than to those of the homogenised one. The bias of the MLE can be eliminated by subsampling at an appropriate rate, which lies between the two characteristic time scales of the problem \cite[Theorems 3.5 and 3.6]{PaS07}. This is similar to the estimation of the integrated stochastic volatility in the presence of market micro-structure noise, where the data has to be subsampled at an appropriate rate \cite{AMZ05,ZMA05}. The correct subsampling rate can be in some instances rather extreme with respect to the frequency of the data and lead to get rid of more than $99\%$ of the data. As the intuition suggests, this increases significantly the bias of the estimator, which is usually taken care of with additional bias corrections and variance reduction procedures.

@@

The rest of the paper is organised as follows. In Section \ref{sec:Setting} we introduce the problem and lay the basis of our analysis setting the main assumptions and notation. In Section \ref{sec:Filter} we present our filtering methodology, with a particular focus on ergodic properties, on multiscale convergence and, naturally, to the properties of our estimators. In Section \ref{sec:Bayesian} we introduce the Bayesian framework and show how it can be enhanced with filtering techniques. Finally, in Section \ref{sec:NumExp} we demonstrate the effectiveness of our methodology via a series of numerical experiments.

@@

In this section, we introduce the class of diffusion processes which we treat in this paper and the classical methodology employed for the estimation of the drift. Let $\epl > 0$ and let us consider the one-dimensional multiscale stochastic differential equation (SDE)
\begin{equation}\label{eq:SDE_MS}
	\d X_t^\epl = -\alpha \cdot V'(X_t^\epl) \dd t - \frac1\epl p'\left(\frac{X_t^\epl}\epl\right) \dd t + \sqrt{2\sigma} \dd W_t,
\end{equation}
where, given a positive integer $N$, we have that $\alpha \in \R^N$ and $\sigma > 0$ are the drift and diffusion coefficients respectively and $W_t$ is a standard one-dimensional Brownian motion. The functions $V\colon \R \to \R^N$ and $p\colon \R \to \R$ correspond to the fast-scale and the slow-scale confining potentials. In particular, we assume 
\begin{equation}\label{eq:Potential}
	V(x) = \begin{pmatrix} V_1(x) & V_2(x) & \cdots & V_N(x) \end{pmatrix}^\top,
\end{equation}
for smooth functions $V_i\colon \R \to \R$, $i = 1, \ldots, N$. Moreover, we assume $p$ to be smooth and periodic of period $L$. Theory of homogenization \cite[Chapter 3]{BLP78} guarantees the existence of an SDE of the form
\begin{equation}\label{eq:SDE_HOM}
	\d X_t = - A \cdot V'(X_t) \dd t + \sqrt{2\Sigma} \dd W_t,
\end{equation}
where $W_t$ is the same Brownian motion as in \eqref{eq:SDE_MS}, such that $X_t^\epl \to X_t$ for $\epl\to 0$ in law as random variables in $\mathcal C^0((0, T); \R)$. In particular, we have $A_i = K\alpha_i$ and $\Sigma = K \sigma$, where the coefficient $K$ is given by the formula
\begin{equation}\label{eq:K_HOM}
	K = \int_0^L (1 + \Phi'(y))^2 \, \mu(\d y),
\end{equation}
with 
\begin{equation}
	\mu(\d y) = \frac1Z e^{-p(y)/\sigma} \dd y, \quad\text{where}\quad Z = \int_0^L e^{-p(y)/\sigma} \dd y,
\end{equation}
and where the function $\Phi$ is the solution of the elliptic partial differential equation
\begin{equation}
	-p'(y)\Phi'(y) + \sigma \Phi''(y) = p'(y), \quad 0 \leq y \leq L,
\end{equation}
endowed with periodic boundary conditions.

We now briefly present the classical methodology for estimating the drift coefficient. Let $T > 0$ and let $X^\epl \defeq (X^\epl_t, 0\leq t, \leq T)$ be a realization of the solution of \eqref{eq:SDE_MS} up to final time. Girsanov's change of measure formula applied to \eqref{eq:SDE_HOM} allows to write the likelihood of $X^\epl$ given a drift coefficient $A$ of \eqref{eq:SDE_HOM} as
\begin{equation}\label{eq:Likelihood}
p(X^\epl \mid A) = \exp\left(-\frac{I(X^\epl\mid A)}{2\Sigma} \right), 
\end{equation}
where 
\begin{equation}
I(X^\epl \mid A) = \int_0^T A \cdot V'(X^\epl_t) \dd X^\epl_t + \frac12 \int_0^T \left( A \cdot V'(X^\epl_t) \right)^2 \dd t.
\end{equation}
Minimizing the functional $I(X^\epl \mid A)$ with respect to $A$ therefore gives the maximum likelihood estimator (MLE) of $A$, which can be formally computed in closed form as
\begin{equation}\label{eq:MLE}
	\widehat A^\epl(T) \defeq \arg \min_{A \in \R^N} I(X^\epl \mid A) = M^{-1}h,
\end{equation}
where $M\in\R^{N\times N}$ and $h\in\R^N$ are defined as
\begin{equation}\label{eq:MandH}
M = \frac1{2\Sigma T} \int_0^T V'(X^\epl_t) \otimes V'(X^\epl_t) \dd t, \quad h = \frac1{2\Sigma T} \int_0^T V'(X^\epl_t) \dd X^\epl_t,
\end{equation}
and where $\otimes$ denotes the tensor product in $\R^N$.
Let us now state the assumptions which will be employed throughout the rest of our work. In particular, we consider the same dissipative setting as \cite[Assumption 3.1]{PaS07}.
\begin{assumption}\label{as:regularity} The potentials $p$ and $V$ satisfy
	\begin{enumerate}
		\item $p \in \mathcal C^\infty(\R) \cap L^\infty(\R)$ and is $L$-periodic for some $L > 0$,
		\item\label{as:regularity_diss} $V_i \in \mathcal C^\infty(\R)$ for all $i=1, \ldots, N$ is polynomially bounded from above and bounded from below, and there exist $a,b > 0$ such that
		\begin{equation}
		-\sum_{i=1}^N V_i'(x)x \leq a - bx^2.
		\end{equation} 
		\item $V'$ is Lipschitz continuous on $\R$: there exists a constant $C > 0$ such that
		\begin{equation}
			\norm{V'(x) - V'(y)} \leq C\abs{x - y},
		\end{equation} 
		\item for all $T > 0$, the symmetric matrix $M$ is positive definite and there exists $\bar \lambda > 0$ such that $\lambda_{\min}(M) \geq \bar \lambda$.
	\end{enumerate}
\end{assumption}

@@

\begin{remark} Given a trajectory $X^\epl \defeq (X^\epl_t, 0\leq t \leq T)$, it is relatively inexpensive to compute $Z^\epl$ from a computational standpoint. In particular, the process $Z^\epl$ is the truncated convolution of the kernel with the process $X^\epl$. Hence, computational tools based on the Fast Fourier Transform (FFT) exist and allow to compute $Z^\epl$ fast component-wise. Moreover, the process $Z_t^\epl$ can be computed, in case $\beta = 1$, in a recursive manner and therefore ``online''.
\end{remark}

In the rest of this section we will focus on the properties of the process $Z^\epl$ when it is considered together with the original process $X^\epl$. In particular, we focus on ergodic properties and multiscale convergence. Finally, we present and analyse an unbiased estimator obtained with the filtering process.

@@

\begin{lemma}\label{lem:ergodicity} Let Assumption \ref{as:regularity} hold and let $b > 0$ be given in Assumption \ref{as:regularity}\ref{as:regularity_diss}. Then, if $\delta > 1/(4b)$, the process $(X^\epl, Z^\epl)^\top$ solution of \eqref{eq:systemSDE} is geometrically ergodic, i.e., there exists $C, \lambda > 0$ such that for all measurable $f$ such that for some integer $q > 0$ 
	\begin{equation}
		f(x) \leq 1 + \abs{x}^q,
	\end{equation}
	it holds
	\begin{equation}
		\abs{\E f(X^\epl_t, Z^\epl_t) - \int_\R\int_\R f(x, z) \rho^\epl(x, z) \dd x \dd z} \leq C\left(1 + \norm{\begin{pmatrix} X^\epl_0 & Z^\epl_0 \end{pmatrix}^\top}^q \right)e^{-\lambda t},
	\end{equation}
	where $\E$ denotes expectation with respect to the Wiener measure and $\rho^\epl$ is the solution to the stationary Fokker--Planck equation
	\begin{equation}
	\label{eq:FPsystem}
	\sigma \partial^2_{xx} \rho^\epl(x,z) +  \partial_x \left( \left( \alpha \cdot V'(x) + \frac{1}{\epl} p' \left ( \frac{x}\epl \right ) \right ) \rho^\epl(x,z) \right) + \frac{1}{\delta} \partial_z \left((z - x) \rho^\epl(x,z)\right) = 0.
	\end{equation}
\end{lemma}

\begin{proof} Lemma \ref{lem:density} guarantees that the Fokker--Planck equation can be written directly from the system \eqref{eq:systemSDE}. For geometric ergodicity, let 
\begin{equation}
	 \mathcal S(x,z) \defeq \begin{pmatrix} -\sum_{i=1}^N V_i'(x) \\ \frac1\delta (x - z) \end{pmatrix} \cdot \begin{pmatrix} x \\ z \end{pmatrix} = -\sum_{i=1}^N V_i'(x)x+\frac1\delta(xz-z^2)
\end{equation}
Due to Assumption \ref{as:regularity}\ref{as:regularity_diss} and Young's inequality, we then have for all $\gamma > 0$
\begin{equation}
	\mathcal S(x,z) \leq a + \left(\frac{1}{2\gamma\delta} - b\right)x^2 + \frac1\delta\left(\frac{\gamma}{2} - 1\right)z^2.
\end{equation}
We choose $\gamma = \gamma^* \defeq 1 - b\delta + \sqrt{2-2b\delta+b^2\delta^2} > 0$ so that
\begin{equation}
	C(\gamma^*) \defeq -\frac{1}{2\gamma^*\delta} + b = -\frac1\delta\left(\frac{\gamma^*}{2} - 1\right),
\end{equation}
and we notice that $C(\gamma^*) > 0$ if $\delta > 1/(4b)$. In this case, we have
\begin{equation}
	\mathcal S(x, z) \leq a - C(\gamma^*) \norm{\begin{pmatrix} x & z \end{pmatrix}^\top}^2,
\end{equation}
and problem \eqref{eq:systemSDE} is dissipative. The result then follows from \cite[Theorem 4.4]{MSH02}.
\end{proof}

\begin{remark} Let us remark that for the multiscale equation \eqref{eq:SDE_MS} one should check the condition above for the drift function $V(x) + \epl^{-1}p(x/\epl)$. Nevertheless, being $p$ bounded, the result holds trivially.  
\end{remark}

\begin{remark} The condition $\delta > 1 / (4b)$ is not very restrictive. Let the parameter dimension $N = 1$ and let $V(x) \propto x^{2p}$ for an integer $p > 1$. Then, Assumption \ref{as:regularity}\ref{as:regularity_diss} holds for an arbitrarily large $b > 0$. Therefore, the filter's parameter $\delta$ can be chosen along the entire positive real axis. A similar argument can be employed for higher dimensions $N > 1$.
\end{remark}

\begin{example}\label{ex:OrnUhl} A closed form solution of \eqref{eq:FPsystem} can be obtained in a simple homogenized case with the dimension of the parameter $N=1$. Let $p(y) = 0$ and the parameters $\alpha$, $\sigma$ be replaced respectively by $A$ and $\Sigma$. Then, if $V(x) = x^2/2$, equation \eqref{eq:FPsystem} has the analytical solution 
\begin{equation}
\rho^0(x,z) = \frac{1}{C_{\rho^0}} \exp\left(-\frac{A}{\Sigma} \frac{x^2}{2} - \frac{1}{\delta \Sigma} \frac{(x - (1+A \delta)z)^2}{2}\right),
\end{equation}
where
\begin{equation}
C_{\rho^0} = \int_{\R} \int_{\R} \exp\left(-\frac{A}{\Sigma} \frac{x^2}{2} - \frac{1}{\delta \Sigma} \frac{(x - (1+A \delta)z)^2}{2}\right) \dd x \dd z.
\end{equation}
This is the density of a multivariate normal distribution $\mathcal N(0, \Gamma)$, where the covariance matrix is given by
\begin{equation}
\Gamma = \frac{\Sigma}{A (1 + A\delta)} \begin{pmatrix} 1+A\delta & 1 \\ 1 & 1 \end{pmatrix}.
\end{equation}
Let us remark that this distribution can be obtained from direct computations involving Gaussian processes. In particular, it is known that $X_t \sim \mathcal{GP}(\mu_t, \mathcal C(t, s))$, where at stationarity $\mu_t = 0$ and
\begin{equation}
	\mathcal C(t, s) = \frac{\Sigma}{A} e^{-A|t-s|}.
\end{equation}
The basic properties of Gaussian processes imply that $Z_t$ is a Gaussian process, and that the couple $(X_t, Z_t)^\top$ is a Gaussian process, too, whose mean and covariance are computable explicitly.
\end{example}

@@

\sigma (\phi^\epl)''(x) + \left ( \alpha \cdot V''(x) + \frac{1}{\epl^2} p'' \left ( \frac{x}\epl \right ) \right ) \phi^\epl(x) + \left ( \alpha \cdot V'(x) + \frac{1}\epl p' \left ( \frac{x}\epl \right ) \right ) (\phi^\epl)'(x) = 0,

@@

Replacing the decomposition \eqref{eq:densityDecomposition} in equation \eqref{eq:FPsystem} and using equation \eqref{FPx} we obtain the following equality

@@

\begin{lemma}\label{lem:convMeasure} Under Assumption \ref{as:regularity}, let $\mu^\epl$ be the invariant measure of the couple $(X^\epl, Z^\epl)^\top$ and let $(X_0^\epl, Z_0^\epl)^\top \sim \mu^\epl$. Then, the measure $\mu^\epl$ converges weakly to the measure $\mu^0(\d x, \d z) = \rho^0(x, z) \dd x \dd z$, whose density $\rho^0$ is the unique solution of the Fokker--Planck equation
\begin{equation} \label{eq:FPsystem_homogenized}
	\Sigma \partial^2_{xx} \rho^0(x,z) + \partial_x\left( A \cdot V'(x) \rho^0(x,z) \right) + \frac{1}{\delta}\partial_z\left((z - x) \rho^0(x,z) \right) = 0,
\end{equation}
where $A$ and $\Sigma$ are the coefficients of the homogenized equation \eqref{eq:SDE_HOM}.
\end{lemma}
\begin{proof} Let $(X, Z)^\top \defeq \left((X_t, Z_t)^\top, 0\leq t \leq T\right)$ be the solution of
	\begin{equation}
	\label{eq:systemSDEHom}
	\begin{aligned}
		\d X_t &= -A \cdot V'(X_t) \dd t + \sqrt{2\Sigma} \dd W_t, \\
		\d Z_t &= \frac{1}{\delta} \left ( X_t - Z_t \right ) \dd t,
	\end{aligned}
	\end{equation} 
	with $(X_0, Z_0)^\top \sim \mu^0$. The arguments of Section \ref{sec:ergodic} can be repeated to conclude that the invariant measure of $(X, Z)^\top$ admits a smooth density $\rho^0$ which satisfies \eqref{eq:FPsystem_homogenized}. Moreover, standard homogenization theory (see e.g. \cite[Chapter 3]{BLP78}) guarantees that$(X^\epl,Z^\epl)^\top \to (X,Z)^\top$ for $\epl \to 0$ in law as random variables with values in $\mathcal C^0((0, T); \R^2)$. The Portmanteau theorem can be employed to conclude that the measure $\mu^\epl$ converges weakly to $\mu^0$ for $\epl \to 0$.
\end{proof}

@@

\begin{proof} Replacing the expression of $\d X^\epl_t$ into \eqref{eq:AHatMixed}, we get for $\widetilde h$

@@

>>>>

\textbf{Abstract.} We study the problem of drift estimation for two-scale continuous time series. We set ourselves in the framework of overdamped Langevin equations, for which a single-scale surrogate homogenized equation exists. In this setting, estimating the drift coefficient of the homogenized equation requires pre-processing of the data, often in the form of subsampling. We avoid subsampling by filtering the data with an appropriate kernel function and compute maximum likelihood estimators based on the filtered process. We show that the estimators we propose are asymptotically unbiased and demonstrate numerically the advantages of our method with respect to subsampling. Finally, we show how our filtering methodology can be combined with Bayesian techniques and provide a full uncertainty quantification of the inference procedure.

@@

Data-driven approaches to the derivation of effective stochastic coarse-grained dynamics have become a standard tool in different areas such as molecular dynamics or atmosphere and ocean science \cite{KKP15,CoP09,YMV19}. When fitting data to a low dimensional coarse-grained model, one is faced with the problem of model misspecification; indeed, the data, coming from the full dynamics, is compatible with the coarse-grained model only at the time scales at which the effective dynamics is valid. This is related to the issue of high-frequency data that arises in econometrics \cite{AiJ14,AMZ05,ZMA05,OSP10}.

The issue of model misspecification in inverse problems with a multiscale structure has been treated in the context of partial differential equations (PDE). In particular, it has been shown that it is possible to infer a coarse-grained equation from data coming from the full model and to retrieve asymptotically the correct result \cite{NPS12}. A series of papers \cite{AbD18, AbD19, AGZ19} focuses on retrieving the full model when the multiscale coefficient is endowed with a specific parametrized structure. Being these problems ill-posed, the latter is achieved via Tikhonov regularization \cite{AbD19,NPS12}, adopting a Bayesian approach \cite{AbD18, NPS12} or exploiting techniques of Kalman filtering \cite{AGZ19}. In \cite{AbD18,AGZ19}, the authors highlight the need of accounting explicitly for the modelling error due to homogenization and apply statistical techniques taken from \cite{CDS18,CES14}.

For simple models in molecular dynamics, the effect of model misspecification was studied in a series of papers \cite{PaS07,PPS09,PPS12,GaS17,GaS18} under the assumption of scale separation. In particular, for Brownian particles moving in two-scale potentials it was shown that, when fitting data from the full dynamics to the homogenized equation, the maximum likelihood estimator (MLE) is asymptotically biased \cite[Theorem 3.4]{PaS07}. To be more precise, in the large sample size limit the MLE converges to the coefficients of the unhomogenized equation, rather than to those of the homogenized one. The bias of the MLE can be eliminated by subsampling at an appropriate rate, which lies between the two characteristic time scales of the problem \cite[Theorems 3.5 and 3.6]{PaS07}. This is similar to the estimation of the integrated stochastic volatility in the presence of market micro-structure noise, where the data has to be subsampled at an appropriate rate \cite{AMZ05,ZMA05}. The correct subsampling rate can be in some instances rather extreme with respect to the frequency of the data and lead to get rid of more than $99\%$ of the data. As the intuition suggests, this increases significantly the bias of the estimator, which is usually taken care of with additional bias corrections and variance reduction procedures.

@@

The rest of the paper is organised as follows. In Section \ref{sec:Setting} we introduce the problem and lay the basis of our analysis setting the main assumptions and notation. In Section \ref{sec:Filter} we present our filtering methodology, with a particular focus on ergodic properties, on multiscale convergence and, naturally, on the properties of our estimators. In Section \ref{sec:Bayesian} we introduce the Bayesian framework and show how it can be enhanced with filtering techniques. Finally, in Section \ref{sec:NumExp} we demonstrate the effectiveness of our methodology via a series of numerical experiments.

@@

In this section, we introduce the class of diffusion processes which we treat in this paper and the classical methodology employed for the estimation of the drift. Let $\epl > 0$ and let us consider the one-dimensional multiscale stochastic differential equation (SDE)
\begin{equation}\label{eq:SDE_MS}
	\d X_t^\epl = -\alpha \cdot V'(X_t^\epl) \dd t - \frac1\epl p'\left(\frac{X_t^\epl}\epl\right) \dd t + \sqrt{2\sigma} \dd W_t,
\end{equation}
where, given a positive integer $N$, we have that $\alpha \in \R^N$ and $\sigma > 0$ are the drift and diffusion coefficients respectively and $W_t$ is a standard one-dimensional Brownian motion. The functions $V\colon \R \to \R^N$ and $p\colon \R \to \R$ correspond to the slow-scale and the fast-scale confining potentials. In particular, we assume 
\begin{equation}\label{eq:Potential}
	V(x) = \begin{pmatrix} V_1(x) & V_2(x) & \cdots & V_N(x) \end{pmatrix}^\top,
\end{equation}
for smooth functions $V_i\colon \R \to \R$, $i = 1, \ldots, N$. Moreover, we assume $p$ to be smooth and periodic of period $L$. Theory of homogenization \cite[Chapter 3]{BLP78} guarantees the existence of an SDE of the form
\begin{equation}\label{eq:SDE_HOM}
	\d X_t = - A \cdot V'(X_t) \dd t + \sqrt{2\Sigma} \dd W_t,
\end{equation}
where $W_t$ is the same Brownian motion as in \eqref{eq:SDE_MS}, such that $X_t^\epl \to X_t$ for $\epl\to 0$ in law as random variables in $\mathcal C^0([0, T]; \R)$. In particular, we have $A = K\alpha$ and $\Sigma = K \sigma$, where the coefficient $K$ is given by the formula
\begin{equation}\label{eq:K_HOM}
	K = \int_0^L (1 + \Phi'(y))^2 \, \mu(\d y),
\end{equation}
with 
\begin{equation}
	\mu(\d y) = \frac1Z e^{-p(y)/\sigma} \dd y, \quad\text{where}\quad Z = \int_0^L e^{-p(y)/\sigma} \dd y,
\end{equation}
and where the function $\Phi$ is the unique solution with zero-mean with respect to the measure $\mu$ of the elliptic partial differential equation
\begin{equation}
	-p'(y)\Phi'(y) + \sigma \Phi''(y) = p'(y), \quad 0 \leq y \leq L,
\end{equation}
endowed with periodic boundary conditions.

We now briefly present the classical methodology for estimating the drift coefficient. Let $T > 0$ and let $X^\epl \defeq (X^\epl_t, 0\leq t \leq T)$ be a realization of the solution of \eqref{eq:SDE_MS} up to final time. Girsanov's change of measure formula applied to \eqref{eq:SDE_HOM} allows to write the likelihood of $X^\epl$ given a drift coefficient $A$ as
\begin{equation}\label{eq:Likelihood}
p(X^\epl \mid A) = \exp\left(-\frac{I(X^\epl\mid A)}{2\Sigma} \right), 
\end{equation}
where 
\begin{equation}
I(X^\epl \mid A) = \int_0^T A \cdot V'(X^\epl_t) \dd X^\epl_t + \frac12 \int_0^T \left( A \cdot V'(X^\epl_t) \right)^2 \dd t.
\end{equation}
Minimizing the functional $I(X^\epl \mid A)$ with respect to $A$ therefore gives the maximum likelihood estimator (MLE) of $A$, which can be formally computed in closed form as
\begin{equation}\label{eq:MLE}
	\widehat A^\epl(T) \defeq \arg \min_{A \in \R^N} I(X^\epl \mid A) = M^{-1}h,
\end{equation}
where $M\in\R^{N\times N}$ and $h\in\R^N$ are defined as
\begin{equation}\label{eq:MandH}
M = \frac1{2\Sigma T} \int_0^T V'(X^\epl_t) \otimes V'(X^\epl_t) \dd t, \quad h = \frac1{2\Sigma T} \int_0^T V'(X^\epl_t) \dd X^\epl_t,
\end{equation}
and where $\otimes$ denotes the outer product in $\R^N$. Let us now state the assumptions which will be employed throughout the rest of our work. In particular, we consider the same dissipative setting as \cite[Assumption 3.1]{PaS07}.
\begin{assumption}\label{as:regularity} The potentials $p$ and $V$ satisfy
	\begin{enumerate}
		\item $p \in \mathcal C^\infty(\R) \cap L^\infty(\R)$ and is $L$-periodic for some $L > 0$,
		\item\label{as:regularity_diss} $V_i \in \mathcal C^\infty(\R)$ for all $i=1, \ldots, N$ is polynomially bounded from above and bounded from below, and there exist $a,b > 0$ such that
		\begin{equation}
		-\sum_{i=1}^N V_i'(x)x \leq a - bx^2.
		\end{equation} 
		\item $V'$ is Lipschitz continuous on $\R$: there exists a constant $C > 0$ such that
		\begin{equation}
			\norm{V'(x) - V'(y)}_2 \leq C\abs{x - y},
		\end{equation} 
		\item for all $T > 0$, the symmetric matrix $M$ is positive definite and there exists $\bar \lambda > 0$ such that $\lambda_{\min}(M) \geq \bar \lambda$.
	\end{enumerate}
\end{assumption}

@@

\begin{remark} Given a trajectory $X^\epl$, it is relatively inexpensive to compute $Z^\epl$ from a computational standpoint. In particular, the process $Z^\epl$ is the truncated convolution of the kernel with the process $X^\epl$. Hence, computational tools based on the Fast Fourier Transform (FFT) exist and allow to compute $Z^\epl$ fast component-wise. Moreover, the process $Z^\epl$ can be computed, in case $\beta = 1$, in a recursive manner and therefore ``online''.
\end{remark}

In the rest of this section we will focus on the properties of the process $Z^\epl$ when it is considered together with the original process $X^\epl$. In particular, we focus on ergodic properties and multiscale convergence. Finally, we present and analyse unbiased estimators obtained with the filtering process.

@@

\begin{lemma}\label{lem:ergodicity} Let Assumption \ref{as:regularity} hold and let $b > 0$ be given in Assumption \ref{as:regularity}\ref{as:regularity_diss}. Then, if $\delta > 1/(4b)$, the process $(X^\epl, Z^\epl)^\top$ solution of \eqref{eq:systemSDE} is geometrically ergodic, i.e., there exists $C, \lambda > 0$ such that for all measurable $f\colon \R^2\to \R$ such that for some integer $q > 0$ 
	\begin{equation}
		f(x, z) \leq 1 + \norm{\begin{pmatrix} x & z \end{pmatrix}^\top}_2^q,
	\end{equation}
	it holds
	\begin{equation}
		\abs{\E f(X^\epl_t, Z^\epl_t) - \int_\R\int_\R f(x, z) \rho^\epl(x, z) \dd x \dd z} \leq C\left(1 + \norm{\begin{pmatrix} X^\epl_0 & Z^\epl_0 \end{pmatrix}^\top}_2^q \right)e^{-\lambda t},
	\end{equation}
	where $\E$ denotes expectation with respect to the Wiener measure and $\rho^\epl$ is the solution to the stationary Fokker--Planck equation
	\begin{equation}
	\label{eq:FPsystem}
	\sigma \partial^2_{xx} \rho^\epl(x,z) +  \partial_x \left( \left( \alpha \cdot V'(x) + \frac{1}{\epl} p' \left ( \frac{x}\epl \right ) \right ) \rho^\epl(x,z) \right) + \frac{1}{\delta} \partial_z \left((z - x) \rho^\epl(x,z)\right) = 0.
	\end{equation}
\end{lemma}

\begin{proof} Lemma \ref{lem:density} guarantees that the Fokker--Planck equation can be written directly from the system \eqref{eq:systemSDE}. For geometric ergodicity, let 
\begin{equation}
	 \mathcal S(x,z) \defeq \begin{pmatrix} -\sum_{i=1}^N V_i'(x) \\ \frac1\delta (x - z) \end{pmatrix} \cdot \begin{pmatrix} x \\ z \end{pmatrix} = -\sum_{i=1}^N V_i'(x)x+\frac1\delta(xz-z^2).
\end{equation}
Due to Assumption \ref{as:regularity}\ref{as:regularity_diss} and Young's inequality, we then have for all $\gamma > 0$
\begin{equation}
	\mathcal S(x,z) \leq a + \left(\frac{1}{2\gamma\delta} - b\right)x^2 + \frac1\delta\left(\frac{\gamma}{2} - 1\right)z^2.
\end{equation}
We choose $\gamma = \gamma^* \defeq 1 - b\delta + \sqrt{1 + (1 - b\delta)^2} > 0$ so that
\begin{equation}
	C(\gamma^*) \defeq -\frac{1}{2\gamma^*\delta} + b = -\frac1\delta\left(\frac{\gamma^*}{2} - 1\right),
\end{equation}
and we notice that $C(\gamma^*) > 0$ if $\delta > 1/(4b)$. In this case, we have
\begin{equation}
	\mathcal S(x, z) \leq a - C(\gamma^*) \norm{\begin{pmatrix} x & z \end{pmatrix}^\top}^2,
\end{equation}
and problem \eqref{eq:systemSDE} is dissipative. The result then follows from \cite[Theorem 4.4]{MSH02}.
\end{proof}

\begin{remark} Let us remark that for the multiscale equation \eqref{eq:SDE_MS} one should check the condition above for the potential $V(x) + p(x/\epl)$. Nevertheless, for functions $p$ satisfying Assumption \ref{as:regularity}, the result holds trivially.  
\end{remark}

\begin{remark} The condition $\delta > 1 / (4b)$ is not very restrictive. Let the parameter dimension $N = 1$ and let $V(x) \propto x^{2p}$ for an integer $p > 1$. Then, Assumption \ref{as:regularity}\ref{as:regularity_diss} holds for an arbitrarily large $b > 0$. Therefore, the parameter of the filter $\delta$ can be chosen along the entire positive real axis. A similar argument can be employed for higher dimensions $N > 1$.
\end{remark}

\begin{example}\label{ex:OrnUhl} A closed form solution of \eqref{eq:FPsystem} can be obtained in a simple homogenized case with the dimension of the parameter $N=1$. Let $p(y) = 0$ and the parameters $\alpha$, $\sigma$ be replaced respectively by $A$ and $\Sigma$. Then, if $V(x) = x^2/2$, equation \eqref{eq:FPsystem} has the analytical solution 
\begin{equation}
\rho^0(x,z) = \frac{1}{C_{\rho^0}} \exp\left(-\frac{A}{\Sigma} \frac{x^2}{2} - \frac{1}{\delta \Sigma} \frac{(x - (1+A \delta)z)^2}{2}\right),
\end{equation}
where
\begin{equation}
C_{\rho^0} = \int_{\R} \int_{\R} \exp\left(-\frac{A}{\Sigma} \frac{x^2}{2} - \frac{1}{\delta \Sigma} \frac{(x - (1+A \delta)z)^2}{2}\right) \dd x \dd z.
\end{equation}
This is the density of a multivariate normal distribution $\mathcal N(0, \Gamma)$, where the covariance matrix is given by
\begin{equation}
\Gamma = \frac{\Sigma}{A (1 + A\delta)} \begin{pmatrix} 1+A\delta & 1 \\ 1 & 1 \end{pmatrix}.
\end{equation}
Let us remark that this distribution can be obtained from direct computations involving Gaussian processes. In particular, it is known that $X \sim \mathcal{GP}(\mu_t, \mathcal C(t, s))$, where at stationarity $\mu_t = 0$ and
\begin{equation}
	\mathcal C(t, s) = \frac{\Sigma}{A} e^{-A|t-s|}.
\end{equation}
The basic properties of Gaussian processes imply that $Z_t$ is a Gaussian process, and that the couple $(X_t, Z_t)^\top$ is a Gaussian process, too, whose mean and covariance are computable explicitly.
\end{example}

@@

\sigma (\phi^\epl)''(x) + \frac{\d}{\d x} \left( \left ( \alpha \cdot V'(x) + \frac{1}{\epl} p' \left ( \frac{x}\epl \right ) \right ) \phi^\epl(x)\right) = 0,

@@

In view of \eqref{eq:densityDecomposition} and \eqref{FPx}, equation \eqref{eq:FPsystem} can be rewritten as

@@

\begin{lemma}\label{lem:convMeasure} Under Assumption \ref{as:regularity}, let $\mu^\epl$ be the invariant measure of the couple $(X^\epl, Z^\epl)^\top$ and let $(X_0^\epl, Z_0^\epl)^\top \sim \mu^\epl$. Then, the measure $\mu^\epl$ converges weakly to the measure $\mu^0(\d x, \d z) = \rho^0(x, z) \dd x \dd z$, whose density $\rho^0$ is the unique solution of the Fokker--Planck equation
\begin{equation} \label{eq:FPsystem_homogenized}
	\Sigma \partial^2_{xx} \rho^0(x,z) + \partial_x\left( A \cdot V'(x) \rho^0(x,z) \right) + \frac{1}{\delta}\partial_z\left((z - x) \rho^0(x,z) \right) = 0,
\end{equation}
where $A$ and $\Sigma$ are the coefficients of the homogenized equation \eqref{eq:SDE_HOM}.
\end{lemma}
\begin{proof} Let $(X, Z)^\top \defeq \left((X_t, Z_t)^\top, 0\leq t \leq T\right)$ be the solution of
	\begin{equation}
	\label{eq:systemSDEHom}
	\begin{aligned}
		\d X_t &= -A \cdot V'(X_t) \dd t + \sqrt{2\Sigma} \dd W_t, \\
		\d Z_t &= \frac{1}{\delta} \left ( X_t - Z_t \right ) \dd t,
	\end{aligned}
	\end{equation} 
	with $(X_0, Z_0)^\top \sim \mu^0$. The arguments of Section \ref{sec:ergodic} can be repeated to conclude that the invariant measure of $(X, Z)^\top$ admits a smooth density $\rho^0$ which satisfies \eqref{eq:FPsystem_homogenized}. Moreover, standard homogenization theory (see e.g. \cite[Chapter 3]{BLP78}) guarantees that $(X^\epl,Z^\epl)^\top \to (X,Z)^\top$ for $\epl \to 0$ in law as random variables with values in $\mathcal C^0([0, T]; \R^2)$. The Portmanteau theorem can be employed to conclude that the measure $\mu^\epl$ converges weakly to $\mu^0$ for $\epl \to 0$.
\end{proof}

@@

\begin{proof} Replacing the expression of $\d X^\epl_t$ into \eqref{eq:MandHTilde}, we get for $\widetilde h$

@@

\begin{equation}\label{eq:alphaDecomposition}
